A Transactional Memory System for Parallel Python

### 2014-08-02 talk/dls2014/paper/paper.tex
some more adjustments

This paper makes the following contributions:
%\vspace{-\topsep}               % XXX:HACK:TODO
\begin{itemize}
\item We introduce a new STM system that performs well for dynamic
  language interpreters.
\vspace{3mm}               % XXX:HACK:TODO
\item We integrate the system closely with a garbage collector
  (GC) to lower the overhead of STM.
\item This new STM system is used to replace the GIL in one
  implementation of Python and is then evaluated.
\end{itemize}

@@

Our TM system is designed as a C library that covers all aspects around
transactions and object memory management. It is designed for
object-oriented dynamic language VMs as a replacement for the GIL.
The library exposes the functions in Listing~\ref{lst:api} as its
interface for use by a dynamic language interpreter.

@@

In the following sections, whenever we use SO, we go through the
address translation to get to the actual contents of an object. This
is also signified by the type \lstinline!object_t!.  This type is
special as it causes the
compiler\footnote{We use Clang 3.5 with patches to deal with bugs in
its ``address-space 256'' feature. Patches are available from authors
until inclusion into the official clang.} to make all accesses through
it relative to the $\%gs$ register.  With exceptions, nearly all
accesses to objects managed by the TM system use this type so
that the CPU will translate the reference to the right version of the
object.
\medskip   % why, Latex, why??

@@

\begin{itemize}
\item prefer transactions that started earlier to younger transactions
  to increase the chance of long transactions succeeding
\item to support \emph{inevitable} transactions, we always prefer them
  to others since they cannot abort (similar to~\cite{blundell06})
\end{itemize}

@@

@@

\section{Related Work}

@@

>>>>

This paper makes the following contributions:
%\vspace{-\topsep}               % XXX:HACK:TODO
\begin{itemize}
\item We introduce a new STM system that performs well for dynamic
  language interpreters.
\vspace{3mm}               % XXX:HACK:TODO
\item We integrate the system closely with a garbage collector
  (GC) to lower the overhead of STM.
\item This new STM system is used to replace the GIL in a Python
  interpreter called PyPy and is then evaluated.
\end{itemize}

@@

Our TM system is designed as a C library~\cite{stmgc-c7} that covers
all aspects around transactions and object memory management. It is
designed for object-oriented dynamic language VMs as a replacement for
the GIL.  The library exposes the functions in Listing~\ref{lst:api}
as its interface for use by a dynamic language interpreter.

@@

In the following sections, whenever we use SO, we go through the
address translation to get to the actual contents of an object. This
is also signified by the type \lstinline!object_t!.  This type is
special as it causes the compiler\footnote{We use Clang 3.5 with
  patches to deal with bugs in its ``address-space 256''
  feature. Patches are available from authors until inclusion into the
  official clang~\cite{stmgc-c7}.} to make all accesses through it
relative to the $\%gs$ register.  With exceptions, nearly all accesses
to objects managed by the TM system use this type so that the CPU will
translate the reference to the right version of the object.
\medskip   % why, Latex, why??

@@

\begin{itemize}
\item prefer transactions that started earlier to younger transactions
  to increase the chance of long transactions succeeding
\item to support \emph{inevitable} transactions, we always prefer them
  to others since they cannot abort~\cite{blundell06}
\end{itemize}

@@

Again we see that the GIL degrades performance on multiple threads.
On the other hand, STM gets faster with more threads, or at least
not much slower. Hence, any application using threads for concurrency
immediately benefits from running on STM. On 4 threads, its performance
is better than the GIL's in 7 out of 8 benchmarks by up to $4\times$.

@@

\vspace{3mm} % formatting hack
\section{Related Work}

@@

\bibitem{stmgc-c7} STMGC-C7 library repository. 2014
  \url{bitbucket.org/pypy/stmgc}
